### 2020-04-19 manuscript/ms.Rmd
add notation to methods

---
title: "Quantifying social distancing behaviour across Australia in response to COVID-19"
author:
  - Nick Golding$^1$, ...$^2$ & ...$^3$
  - $^1$ Telethon Kids Institute and Curtin University, Perth WA, Australia
  - $^2$ ...
  - $^3$ ...
bibliography: distancing.bib
output:
  pdf_document:
  word_document:
    fig_caption: true
    highlight: null
header-includes: \usepackage{float}
---

## Gist (to be split up into a paper format)

Social distancing behaviour - advised and enforced by governments around the world - has been a cornerstone of the global response to the COVID-19 pandemic. Quantifying how the behaviour of a population has changed in response to these public health measures, is critical to monitoring the ongoing effectiveness of these measures, identifying which are most impactful, and as inputs to modelling to forecast how the measures will aimpact ont he course of the outbreak.

A variety of data sources have are available that can quantify different aspects of social distancing behaviour at a population level. Aggregated user-tracking data from technology companies provide timeseries of: how often users leave home, the lengths of journeys they undertake when they do, the types of locations they visited, and the locations of those places.

Whilst these multiple data sources provide varied perspectives on population behaviour, they vary significantly in the subset of the population they reflect and the aspects of behaviour being considered. For example, data from route-planning apps is more likely to reflect unfamiliar journeys than trips to places regularly visited, and data on the types of places being visited does not indicate how far users travel to visit them. The different, and often non-overlapping spatial scales and time frames of these data sources also hinders direct comparison and synthesis of them.

Moreover, no single one of these types of data fully encompasses the range of behaviours that affect transmission of COVID-19 - those that can be sumamrised in a mathematical modelling framework via the mixing parameter of a transmission model.

The aim of this work is to synthesise these varied measures of population mixing behaviour into a small, quantified set of latent social distancing factors that reflect the underlying changes in behaviour. We consider these latent social distancing factors in a semi-mechanistic way; reflecting knowledge and hypotheses of the events that triggered and influenced these behavioural trends. To do this, we construct a statistical semi-mechanistic latent factor model that considers dates that events, policy measures, and public announcements are made, and estimates how the behaviour of the population responded to each events by calibration against all of the available metrics of population mixing behaviour.

We apply this model to 9 separate data streams provided by from two technology companies for Australia, with spatial data varying between national, state and city-levels. This reveals two distinct latent population behavioural factors: a strong social distancing effect in response to national announcements and policies, preceeded by a small and short-lived rise in mixing apparently driven by the population 'stocking up' on supplies.

To add:
 - relative effectiveness of three different policies
 - relative effect size by state

## Abstract

 - Social distancing 
 - Data available
 - Synthesis needed
 - Findings
 - Use for monitoring

## Introduction

 - COVID-19 background
 - Social distancing policies
 - Population mixing/mobility studies
 - Timeline of data availability

## Methods

### General model description

#### Latent factor model

We construct a mechanistic latent-factor model that relates a series of underlying ('latent') population behaviours to an array of data streams, and then infer its parameters using statistical inference. This section, gives a general formulation of the model. Details of the specific implementation and data sources for Australia are given in a later section.

For each data stream $i$, we observe a time series of data values $\mathbf{y}_{i}$, comprising the values of the metric at each time $t$. We model the overall trend $\mathbf{m}_i$ in each of these data streams as a function of $K$ different temporal latent factors $\mathbf{z}_k$.

Each datastream trend is modelled as a linear function of the latent behavioural factors, with intercept term $\alpha_i$ and loading values $\omega_{i,k}$ against each latent factor:

$$
m_{i,t} = \alpha_i + \sum_{k=1}^K \omega_{i,k} z_{k,t}
$$
The latent factors $\mathbf{z}_k$ quantify general patterns of behaviour, rather than any one specific measurement of behaviour. As such, they are unitless, relative values which for statistical convenience we constrain to be scaled between 0 and 1. The sign and magnitude of the loadings $\omega_{i,k}$ therefore determine whether each behaviour corresponds to an increase or decrease in that metric, and by what magnitude.

#### Mechanistic latent factors

  We model each of these latent behavioural factors in a semi-mechanistic fashion. We compute the degree of behavioural change as a function of the timing, magnitude and early/late adoption of behavioural responses to events. For each latent factor $k$, we model the impact of each event $j$ as a function of: $\tau_{j,k}$ the time when the event took place; $\epsilon_{j,k}$ the relative impact of the event on behaviour; and $\kappa_{j,k}$, the variability among the population in how rapidly individuals responded to these events.

  For each latent behaviour $k$, these parameters describe the behavioural reponse to each event via a logistic distribution. Where the behaviour occurs at a single point in time, these parameters give rise to a logistic distribution over the times at which members of the population exhibit this behaviour, where the fraction of the population exhibiting behaviour $k$ in response to event $j$ at time $t$ is given by the probability density function of the logistic distribution:

$$
p_{j,k}(t) = \frac{e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}{\kappa_{j,k} (1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}) ^ 2}
$$

  Where the latent behavioural state is described by individuals switching from one behavioural state to another, the fraction of the population that have made the switch by time $t$ is given by the cumulative density function of the logistic distribution:

$$
p_{j,k}(t) = \frac{1}{1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}
$$

  Since the latent factors are scaled between 0 and 1, we scale these probabilities such that their maximum value is the relative event impact parameter $\epsilon_{j,k}$, and the sum over these parameeters is constrained to 1. Therefore, where the behaviour is a function of only a single event, $\epsilon_{j,k}$ is fixed at 1. The contribution of each event to the behaviour is therefore given by:

<!-- MAKE THIS HAVE DENSITY EPSILON AT THE MEAN -->
<!-- HAS DENSITY 1/(4 * kappa) AT MEAN -->
<!-- SO DIVIDE BY THAT AND MULTIPLY BY EPSILON -->
<!-- MUTIPLY BY epsilon * 4 * kappa -->

<!-- CHECK THIS! -->
$$
f_{j,k}(t) = 4 \kappa_{j,k} \epsilon_{j,k} \frac{e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}{\kappa_{j,k} (1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}) ^ 2}
$$
And the cumulative contribution by:

$$
f_{j,k}(t) = \frac{\epsilon_{j,k}}{1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}
$$

#### Data observatrion models

   - observation model (weekends, holidays, and interactions)

 - Data availability by source
 - Announcement timeline for Australia
 - Specific model fitting for Australia
    - interaction between distancing and weekends
    - greta HMC
    - residuals checking

## Results

Figures:
 - Two latent factors nationally
 - Latent factors superimposed on different data sources
 - Loadings for each data type (ordination-style plot of similarities between them? After taking absolute value)
 - Relative impact of different announcements (summary of ranking in posterior?)
 - Variation in latent factors by state

## Discussion

 - Use for monitoring adherence.
 - Use in modelling studies.
 - Caveat that the population is biased to users of these technologies.
 - Potential to include more data sources.