### 2021-12-09 Rmd/case_date_autocorrelations.Rmd
autocorrelation of cases by onset and confirmation dates

```{r}
nndss_ll <- get_nndss_linelist()
vic_ll <- get_vic_linelist(file = "~/not_synced/vic/20211206_linelist_reff.csv")
nsw_ll <- get_nsw_linelist()

vl <- vic_ll %>% filter(import_status == "local", date_confirmation >= "2021-08-01")
vn <- nndss_ll %>% filter(import_status == "local", date_confirmation >= "2021-08-01", state == "VIC")

nl <- nsw_ll %>% filter(import_status == "local", date_confirmation >= "2021-08-01")
nn <- nndss_ll %>% filter(import_status == "local", date_confirmation >= "2021-08-01", state == "NSW")

dat <- bind_rows(
  vl %>% mutate(linelist = "state"),
  vn %>% mutate(linelist = "NINDSS"),
  nl %>% mutate(linelist = "state"),
  nn %>% mutate(linelist = "NINDSS"),
)
```

Confirmation dates
```{r}
dat %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_confirmation,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  )
```

Confirmation dates last month
```{r}
dat %>%
  filter(date_confirmation >= "2021-11-01") %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_confirmation,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  )
```

Onset dates
```{r}
dat %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_onset,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  ) +
  coord_cartesian(ylim = c(0,1))
```

```{r}
dat %>%
  ggplot() +
  geom_point(
    aes(x = date_onset, y = date_confirmation)
  ) +
  facet_grid(rows = vars(state), cols = vars(linelist)) +
  scale_x_date(date_breaks = "month") +
  theme(axis.text.x = element_text(angle = 270))

```

```{r}
dat %>%
  filter(date_onset >= "2021-08-01") %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_onset,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  ) 
```

Onset dates last month
```{r}
dat %>%
  filter(date_onset >= "2021-11-01") %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_onset,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  )
```

Detection dates
```{r}
dat %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_detection,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  ) +
  coord_cartesian(ylim = c(0,1))
```

Onset dates last month
```{r}
dat %>%
  filter(date_detection >= "2021-11-01") %>%
  ggplot() +
  geom_bar(
    aes(
      x = date_detection,
      fill = linelist
    ),
    position = "dodge"
  ) +
  facet_grid(
    rows = vars(state)
  )
```

```{r}
dates <- expand_grid(
  date = seq.Date(from = as.Date("2021-08-01"), to = as.Date("2021-12-05"), by = 1),
  state = c("NSW", "VIC"),
  linelist = c("state", "NINDSS")
)

dat_conf <- dat %>%
  group_by(state, linelist, date_confirmation) %>%
  summarise(cases = n(), .groups = "drop") %>%
  filter(!is.na(date_confirmation)) %>%
  full_join(
    dates %>% rename(date_confirmation = date)
  ) %>%
  arrange(state, linelist, date_confirmation) %>%
  mutate(cases = ifelse(is.na(cases), 0, cases))

dat_onset <- dat %>%
  group_by(state, linelist, date_onset) %>%
  summarise(cases = n(), .groups = "drop") %>%
  filter(!is.na(date_onset)) %>%
  full_join(
    dates %>% rename(date_onset = date)
  ) %>%
  arrange(state, linelist, date_onset) %>%
  mutate(cases = ifelse(is.na(cases), 0, cases))

dcn <- dat_conf %>% filter(state == "NSW", linelist == "state", date_confirmation >= "2021-10-01")
dcv <- dat_conf %>% filter(state == "VIC", linelist == "state", date_confirmation >= "2021-10-01")

don <- dat_onset %>% filter(state == "NSW", linelist == "state", date_onset >= "2021-10-01")
dov <- dat_onset %>% filter(state == "VIC", linelist == "state", date_onset >= "2021-10-01")
```

```{r}
library(tseries)
```

Autocorrelation

VIC

confirmation dates

since oct
```{r}
acf(dcv$cases)
```

Since Nov
```{r}
dcv %>% filter(date_confirmation >= "2021-11-01") %>% pull(cases) %>% acf
```

Onset dates

since Oct
```{r}
acf(dov$cases)
```

Since Nov
```{r}
dov %>% filter(date_onset >= "2021-11-01") %>% pull(cases) %>% acf
```

NSW

Confrimation dates 
since oct

```{r}
acf(dcn$cases)
```

since nov
```{r}
dcn %>% filter(date_confirmation >= "2021-11-01") %>% pull(cases) %>% acf
```

onset dates

since oct
```{r}
acf(don$cases)
```

since nov
```{r}
don %>% filter(date_onset >= "2021-11-01") %>% pull(cases) %>% acf
```